---
title: "imageScrambleR-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{imageScrambleR-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Once _imageScrambleR_ is installed, load the package.
```{r setup, echo=TRUE}
library(imageScrambleR)
```

To demonstrate _imageScrambleR_'s current capabilities, the following code loads an image using the _magick_ package, then displays it using that package's print method.
```{r loadImage}
r_logo_path <- file.path(R.home("doc"), "html", "logo.jpg")
library(magick)
r_logo <- image_read(r_logo_path)
print(r_logo)
```

The _magick2arr_ function is used to convert these images to grayscale, then to matrices with hexadecimal values for each pixel.
```{r m2r}
r_logo_arr <- magick2arr(r_logo)
head(r_logo_arr)
```

We can then use any of the four image scrambler functions to scramble the image using an integer to indicate how many chunks we want each dimension split into.
```{r scram}
r_logo_arr_scram <- scrambler_naiveR(r_logo_arr, 3L)
r_logo_arr_scram <- scrambler_R(r_logo_arr, 3L)
r_logo_arr_scram <- scrambler_Cpp(r_logo_arr, 3L)
r_logo_arr_scram <- scrambler_Cpp2(r_logo_arr, 3L)
```

Now we can use the _arr2magick_ function to convert the image array back to _magick_ image format, and view it.
```{r}
r_logo_scram <- arr2magick(r_logo_arr_scram)
print(r_logo_scram)
```

The R package _microbenchmark_ can be used to run the four image scramblers 100 times each, using the same input parameters.
Output time is in nanoseconds.
```{r benchmark_r, echo=TRUE}
benchmarks_r_logo <- microbenchmark::microbenchmark(
    scrambler_R(r_logo_arr, 3L),
    scrambler_naiveR(r_logo_arr, 3L),
    scrambler_Cpp2(r_logo_arr, 3L),
    scrambler_Cpp(r_logo_arr, 3L))
print(benchmarks_r_logo)
plot(benchmarks_r_logo, names = c('R', 'naiveR', 'Cpp2', 'Cpp'))
```
